Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices

### 2025-03-03 jso.qmd
all to US spelling

---
title: Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices
author:
  - name: H. Sherry Zhang
    email: huize.zhang@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Austin
          country: United States
          postal-code: 78751
    attributes:
        corresponding: true
  - name: Dianne Cook
    email: dicook@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
  - name: Nicolas  Langrené
    email: nicolaslangrene@uic.edu.cn
    affiliations:
        - id: 3
          name: Guangdong Provincial/Zhuhai Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
          department: Department of Mathematical Sciences
          city: Zhuhai
          country: China
          postal-code: 519087
  - name: Jessica Wai Yin Leung
    email: Jessica.Leung@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
abstract: |
  The projection pursuit (PP) guided tour interactively optimises a criterion function known as the PP index, to explore high-dimensional data by revealing interesting projections. Optimisation of some PP indexes can be non-trivial, if they are non-smooth functions, or the optimum has a small "squint angle", detectable only from close proximity. Here, measures for calculating the smoothness and squintability properties of the PP index are defined. These are used to investigate the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes. The performance of JSO for visualising data is evaluated across various hyper-parameter settings and compared with existing optimisers for the guided tour. The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability measures are implemented in the `ferrn` package. 
keywords: 
  - projection pursuit
  - jellyfish search optimiser (JSO)
  - optimisation
  - grand tour
  - high-dimensional data
  - exploratory data analysis
date: last-modified
bibliography: bibliography.bib
format:
  tandf-pdf:
    keep-tex: true
    fontsize: 12pt
    linestretch: 2
editor: 
  markdown: 
    wrap: 72
crossref: 
  eq-prefix: ""
editor_options: 
  chunk_output_type: console
header-includes:
   - \usepackage{algorithm}
   - \usepackage{float} 
   - \usepackage{amsmath}
   - \usepackage{amsthm}
   - \usepackage{amssymb}
   - \theoremstyle{plain}
   - \newtheorem{defn}{\protect\definitionname}
   - \newtheorem{prop}{\protect\propositionname}
   - \providecommand{\definitionname}{Definition}
   - \providecommand{\propositionname}{Proposition}
nocite: |
   @DLMF
---

@@

Projection pursuit (PP) [@kr69; @FT74; @huber85] is a dimension
reduction technique aimed at identifying informative linear projections
of data. This is useful for exploring high-dimensional data, and
creating plots of the data that reveal the main features to use for
publication. The method involves optimising an objective function known
as the PP index [e.g., @hall1989polynomial; @cook1993projection;
@lee2010projection; @Loperfido2018; @Loperfido2020], which defines the
criterion for what constitutes interesting or informative projections.
Let $X \in \mathbb{R}^{n\times p}$ be the data matrix,
$A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$
belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The
projection $Y = XA$ is a linear transformation that maps data from a
$p$-dimensional space into a $d$-dimensional space. The index function
$f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function
that measures an interesting aspect of the projected data, such as
deviation from normality, presence of clusters, or non-linear structure.
For a fixed sample of data, PP finds the orthonormal basis $A$ that
maximises the index value of the projection, $Y = XA$:

@@

Optimisation of PP is often discussed when new indexes are proposed
[@posse95; @marie-sainte2010; @grochowski2011]. @cook1995grand tied the
optimisation more closely to the index, when they introduced the PPGT,
which monitors the optimisation visually so that the user can see the
projected data leading in and out of the optimum. An implementation is
available in the `tourr` package [@tourr] in R [@R]. @RJ-2021-105
illustrated how to diagnose optimisation processes, particularly
focusing on the guided tour, and revealed a need for improved
optimisation. While improving the quality of the optimisation solutions
in the tour is essential, it is also important to be able to view the
data projections as the optimisation progresses. Integrating the guided
tour with a global optimisation algorithm that is efficient in finding
the global optimal and enables viewing of the projected data during the
exploration process is a goal.

Here, the potential for a Jellyfish Search Optimiser (JSO) [see
@chou_novel_2021, and @rajwar_exhaustive_2023] for the PPGT is explored.
JSO, inspired by the search behaviour of jellyfish in the ocean, is a
swarm-based metaheuristic designed to solve global optimisation
problems. Compared to traditional methods, JSO has demonstrated stronger
search ability and faster convergence, and requires fewer tuning
parameters. These practicalities make JSO a promising candidate for
enhancing PP optimisation.

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimisation for the guided tour. It is of
interest to assess how quickly and closely the optimiser reaches a
global optimum, for various PP indexes that may have differing
complexities. To observe the performance of JSO with different types of
PP indexes, metrics are introduced to capture specific properties of the
index including squintability [based on @barnett1981interpreting's
squint angle] and smoothness. Here, we mathematically define metrics for
squintability and smoothness, which is a new contribution for PP
research. A series of simulation experiments using various datasets and
PP indexes are conducted to assess JSO's behaviour and its sensitivity
to hyper-parameter choices (number of jellyfish and maximum number of
tries). The relationship between the JSO performance, hyper-parameter
choices and properties of PP indexes (smoothness and squintability) is
analysed to provide guidance on selecting optimisers for practitioners
using projection pursuit. Additionally, this work should guide the
design of new PP indexes and facilitate better optimization for PP.

@@

-   a pseudo-derivative, that searches locally for the best direction,
    based on differencing the index values for very close projections.
-   a brute-force optimisation (CRS).
-   a modified brute force algorithm described in @posse95.
-   an essentially simulated annealing [@Bertsimas93] where the search
    space is reduced during the optimisation.
-   a very localised search, to take tiny steps to get closer to the
    local maximum.

```{=html}
<!-- 
-   `search_geodesic()`: provides a pseudo-derivative optimisation. It
    searches locally for the best direction, based on differencing the
    index values for very close projections. Then it follows the
    direction along the geodesic path between planes, stopping when the
    next index value fails to increase.
-   `search_better()`: also known as Creeping Random Search (CRS), is a
    brute-force optimisation searching randomly for projections with
    higher index values.
-   `search_better_random()`: is essentially simulated annealing
    [@Bertsimas93] where the search space is reduced as the optimisation
    progresses.
-   `search_posse()`: implements the algorithm described in @posse95.
-   `search_polish()`: is a very localised search, to take tiny steps to
    get closer to the local maximum.
-->
```
There are numerous PP index functions available: introduced in @huber85,
@cook1993projection, @lee2005projection, @lee2010projection, @Grimm2016,
@Laa:2020wkm. Most are relatively simply defined, for any projection
dimension, and implemented because they are relatively easy to optimise.
A goal is to develop PP indexes based on scagnostics [@scag, @WW08], but
the blockage is their optimisation as these tend to be noisy, with
potentially small squint angles.

@@

The Jellyfish Search Optimiser (JSO) mimics the natural movements of
jellyfish, which include passive and active motions driven by ocean
currents and their swimming patterns, respectively. In the context of
optimization, these movements are abstracted to explore the search
space, aiming to balance exploration (searching new areas) and
exploitation (focusing on promising areas). The algorithm aims to find
the optimal solution by adapting the behaviour of jellyfish to navigate
towards the best solution over iterations [@chou_novel_2021].

@@

The performance of the optmisers is measured by the success rate, which is defined as the proportion of simulations that achieves a final index value within 0.05
of the best index value found among all 50 simulations. @fig-success-rate illustrates why the choice of 0.05 is reasonable: the pipe is not recognisable in the projected data. This is motivated by @laa_using_2020's approach to investigating PP indexes.

@@

A generalised linear model is fitted using a quasibinomial family and a logit link function to assess the factors affecting success rate.  Predictors are smoothness,
squintability, data dimension and JSO hyper-parameters. <!-- The data is
pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
for interpretation, 2) creating a new binary variable `long_time` to
indicate cases with an average run time over 20 seconds, and 3)
re-coding the success rate for the `stringy` index as 0, because none of
the 50 simulations correctly identified the sine wave shape.-->

@@

The performance of JSO is assessed relative to the existing CRS optimiser based on the optimal projection obtained. The final projections for the pipe data sets found by the two optimisers are shown in @fig-proj, with rows corresponding to data dimension. The 50 simulations in each data dimension are sorted by index value, and the projections corresponding to each 10th quantile value are shown, with minimum at left and maximum at right. Index value is printed on the plot. The purpose is to summarise the views of the data resulting from different optimisations, and hence compare results between the two optimisers. Generally, the JSO does a more consistent job of finding the pipe structure clearly. As the data dimension increases, both optimisers struggle and less clearly capture the circle shape.

@@

The effect of JSO hyper-parameters (number of jellyfish and the maximum
number of tries) on the success rate is summarised in @fig-proportion.
The uncertainty is quantified through 500 bootstrap replicates for each
case. As the number of jellyfish and maximum tries increase, the success
rate also increases. For problems with relatively lower dimensional (4, 6) search spaces, small parameter values (20 jellyfish and a maximum of 50 tries) are sufficient to achieve a high success rate. Many more jellyfish (100 jellyfish and a maximum of 100 tries) are needed to obtain a mediocre success rate for dimensions 10 and 12. This suggests that even JSO will struggle with high dimensions, and require substantial more computational time to have an acceptable success rate, especially for difficult PP indexes based on scagnostics. <!-- dimensions) where a more complex search is necessary. However, in lower-dimensional problems, the benefit may be less pronounced, as the space is already relatively easy to navigate. While increasing both parameters enhances the performance of JSO, it also extends the computational time required for the optimisation, which can be
computationally intensive when evaluating the index function (such as
scagnostic indexes) multiple times across numerous iterations.-->

@@

@fig-idx-proj-dist shows a sample of traces from the indexes in @tbl-smoothness-squintability for the 6D sine wave data. The traces the PP index values computed for projections along an interpolated path between a noise projection and the optimal projection (sine wave). (Note index values were self-standardised to range between 0 and 1, for comparison.) The `splines` index is very smooth with a gradual increase in value between the noise projection and the optimal projection. The `TIC` index is slightly less smooth and slightly less gradual change in value. Least desirable is the `skinny` index which is noisier and the optimisation needs to be much closer to the optimal angle to have a higher value. This illustrates how the new measures for squintability and smoothness work to describe the PP indexes.

@@

The simulation studies have compared the JSO with CRS and shown how
smoothness and squintability affect the JSO success rate. Here we
explain how they can be used for new index and optimiser development.
Using the JSO optimiser for PPGT requires a change in user behaviour.
For all the existing optimisers a single optimisation path is followed.
The JSO has multiple optimisation paths, and needs additional tools to
incorporate into the PPGT, as explained here.

@@

To use JSO for PPGT in the `tourr` package, specify
`search_f = search_jellyfish` in the guided tour. Unlike existing
optimisers, the animation function `animate_*()` won't directly render
the tour path for JSO due to the generation of multiple tour paths. To
visualise the path visited by each individual jellyfish, assign the
animation to an object (`res <- animate_xy(...)`). This will save the
bases visited by JSO, along with relevant metadata, as a tibble data
object [see @RJ-2021-105 for more details on the data object]. The bases
visited for individual jellyfish can then be extracted and viewed using
the `planned_tour()` function.

@@

To sample bases for calculating smoothness, the following arguments are
required: the index function, the dataset, and the number of random
bases to sample. The output of `sample_bases()` is a data frame with a
list-column of sampled basis matrix and index value. Parallellasation is
available to speed up the index value computation through the
`parallel = TRUE` argument.

@@

This paper has presented new metrics to mathematically define desirable
features of PP indexes, squintability and smoothness, and used these to
assess the performance of the new jellyfish search optimiser. The
metrics will be generally useful for characterising PP indexes, and help
with developing new indexes.

>>>>

---
title: Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices
author:
  - name: H. Sherry Zhang
    email: huize.zhang@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Austin
          country: United States
          postal-code: 78751
    attributes:
        corresponding: true
  - name: Dianne Cook
    email: dicook@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
  - name: Nicolas  Langrené
    email: nicolaslangrene@uic.edu.cn
    affiliations:
        - id: 3
          name: Guangdong Provincial/Zhuhai Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
          department: Department of Mathematical Sciences
          city: Zhuhai
          country: China
          postal-code: 519087
  - name: Jessica Wai Yin Leung
    email: Jessica.Leung@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
abstract: |
  The projection pursuit (PP) guided tour interactively optimizes a criterion function known as the PP index, to explore high-dimensional data by revealing interesting projections. Optimization of some PP indexes can be non-trivial, if they are non-smooth functions, or the optimum has a small "squint angle", detectable only from close proximity. Here, measures for calculating the smoothness and squintability properties of the PP index are defined. These are used to investigate the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimizer (JSO), for optimizing PP indexes. The performance of JSO for visualizing data is evaluated across various hyper-parameter settings and compared with existing optimisers for the guided tour. The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability measures are implemented in the `ferrn` package.
keywords: 
  - projection pursuit
  - jellyfish search optimiser (JSO)
  - optimisation
  - grand tour
  - high-dimensional data
  - exploratory data analysis
date: last-modified
bibliography: bibliography.bib
format:
  tandf-pdf:
    keep-tex: true
    fontsize: 12pt
    linestretch: 2
editor: 
  markdown: 
    wrap: 72
crossref: 
  eq-prefix: ""
editor_options: 
  chunk_output_type: console
header-includes:
   - \usepackage{algorithm}
   - \usepackage{float} 
   - \usepackage{amsmath}
   - \usepackage{amsthm}
   - \usepackage{amssymb}
   - \theoremstyle{plain}
   - \newtheorem{defn}{\protect\definitionname}
   - \newtheorem{prop}{\protect\propositionname}
   - \providecommand{\definitionname}{Definition}
   - \providecommand{\propositionname}{Proposition}
nocite: |
   @DLMF
---

@@

Projection pursuit (PP) [@kr69; @FT74; @huber85] is a dimension
reduction technique aimed at identifying informative linear projections
of data. This is useful for exploring high-dimensional data, and
creating plots of the data that reveal the main features to use for
publication. The method involves optimizing an objective function known
as the PP index [e.g., @hall1989polynomial; @cook1993projection;
@lee2010projection; @Loperfido2018; @Loperfido2020], which defines the
criterion for what constitutes interesting or informative projections.
Let $X \in \mathbb{R}^{n\times p}$ be the data matrix,
$A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$
belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The
projection $Y = XA$ is a linear transformation that maps data from a
$p$-dimensional space into a $d$-dimensional space. The index function
$f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function
that measures an interesting aspect of the projected data, such as
deviation from normality, presence of clusters, or non-linear structure.
For a fixed sample of data, PP finds the orthonormal basis $A$ that
maximizes the index value of the projection, $Y = XA$:

@@

Optimization of PP is often discussed when new indexes are proposed
[@posse95; @marie-sainte2010; @grochowski2011]. @cook1995grand tied the
optimisation more closely to the index, when they introduced the PPGT,
which monitors the optimisation visually so that the user can see the
projected data leading in and out of the optimum. An implementation is
available in the `tourr` package [@tourr] in R [@R]. @RJ-2021-105
illustrated how to diagnose optimisation processes, particularly
focusing on the guided tour, and revealed a need for improved
optimisation. While improving the quality of the optimisation solutions
in the tour is essential, it is also important to be able to view the
data projections as the optimisation progresses. Integrating the guided
tour with a global optimisation algorithm that is efficient in finding
the global optimal and enables viewing of the projected data during the
exploration process is a goal.

Here, the potential for a Jellyfish Search Optimizer (JSO) [see
@chou_novel_2021, and @rajwar_exhaustive_2023] for the PPGT is explored.
JSO, inspired by the search behavior of jellyfish in the ocean, is a
swarm-based metaheuristic designed to solve global optimisation
problems. Compared to traditional methods, JSO has demonstrated stronger
search ability and faster convergence, and requires fewer tuning
parameters. These practicalities make JSO a promising candidate for
enhancing PP optimisation.

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimisation for the guided tour. It is of
interest to assess how quickly and closely the optimiser reaches a
global optimum, for various PP indexes that may have differing
complexities. To observe the performance of JSO with different types of
PP indexes, metrics are introduced to capture specific properties of the
index including squintability [based on @barnett1981interpreting's
squint angle] and smoothness. Here, we mathematically define metrics for
squintability and smoothness, which is a new contribution for PP
research. A series of simulation experiments using various datasets and
PP indexes are conducted to assess JSO's behavior and its sensitivity
to hyper-parameter choices (number of jellyfish and maximum number of
tries). The relationship between the JSO performance, hyper-parameter
choices and properties of PP indexes (smoothness and squintability) is
analysed to provide guidance on selecting optimisers for practitioners
using projection pursuit. Additionally, this work should guide the
design of new PP indexes and facilitate better optimization for PP.

@@

-   a pseudo-derivative, that searches locally for the best direction,
    based on differencing the index values for very close projections.
-   a brute-force optimisation (CRS).
-   a modified brute force algorithm described in @posse95.
-   an essentially simulated annealing [@Bertsimas93] where the search
    space is reduced during the optimisation.
-   a very localized search, to take tiny steps to get closer to the
    local maximum.

```{=html}
<!-- 
-   `search_geodesic()`: provides a pseudo-derivative optimisation. It
    searches locally for the best direction, based on differencing the
    index values for very close projections. Then it follows the
    direction along the geodesic path between planes, stopping when the
    next index value fails to increase.
-   `search_better()`: also known as Creeping Random Search (CRS), is a
    brute-force optimisation searching randomly for projections with
    higher index values.
-   `search_better_random()`: is essentially simulated annealing
    [@Bertsimas93] where the search space is reduced as the optimisation
    progresses.
-   `search_posse()`: implements the algorithm described in @posse95.
-   `search_polish()`: is a very localised search, to take tiny steps to
    get closer to the local maximum.
-->
```
There are numerous PP index functions available: introduced in @huber85,
@cook1993projection, @lee2005projection, @lee2010projection, @Grimm2016,
@Laa:2020wkm. Most are relatively simply defined, for any projection
dimension, and implemented because they are relatively easy to optimist.
A goal is to develop PP indexes based on scagnostics [@scag, @WW08], but
the blockage is their optimisation as these tend to be noisy, with
potentially small squint angles.

@@

The Jellyfish Search Optimizer (JSO) mimics the natural movements of
jellyfish, which include passive and active motions driven by ocean
currents and their swimming patterns, respectively. In the context of
optimization, these movements are abstracted to explore the search
space, aiming to balance exploration (searching new areas) and
exploitation (focusing on promising areas). The algorithm aims to find
the optimal solution by adapting the behavior of jellyfish to navigate
towards the best solution over iterations [@chou_novel_2021].

@@

The performance of the opt misers is measured by the success rate, which is defined as the proportion of simulations that achieves a final index value within 0.05
of the best index value found among all 50 simulations. @fig-success-rate illustrates why the choice of 0.05 is reasonable: the pipe is not recognizable in the projected data. This is motivated by @laa_using_2020's approach to investigating PP indexes.

@@

A generalized linear model is fitted using a quasibinomial family and a logit link function to assess the factors affecting success rate.  Predictors are smoothness,
squintability, data dimension and JSO hyper-parameters. <!-- The data is
pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
for interpretation, 2) creating a new binary variable `long_time` to
indicate cases with an average run time over 20 seconds, and 3)
re-coding the success rate for the `stringy` index as 0, because none of
the 50 simulations correctly identified the sine wave shape.-->

@@

The performance of JSO is assessed relative to the existing CRS optimiser based on the optimal projection obtained. The final projections for the pipe data sets found by the two optimisers are shown in @fig-proj, with rows corresponding to data dimension. The 50 simulations in each data dimension are sorted by index value, and the projections corresponding to each 10th quantile value are shown, with minimum at left and maximum at right. Index value is printed on the plot. The purpose is to summarize the views of the data resulting from different optimization, and hence compare results between the two optimisers. Generally, the JSO does a more consistent job of finding the pipe structure clearly. As the data dimension increases, both optimisers struggle and less clearly capture the circle shape.

@@

The effect of JSO hyper-parameters (number of jellyfish and the maximum
number of tries) on the success rate is summarized in @fig-proportion.
The uncertainty is quantified through 500 bootstrap replicates for each
case. As the number of jellyfish and maximum tries increase, the success
rate also increases. For problems with relatively lower dimensional (4, 6) search spaces, small parameter values (20 jellyfish and a maximum of 50 tries) are sufficient to achieve a high success rate. Many more jellyfish (100 jellyfish and a maximum of 100 tries) are needed to obtain a mediocre success rate for dimensions 10 and 12. This suggests that even JSO will struggle with high dimensions, and require substantial more computational time to have an acceptable success rate, especially for difficult PP indexes based on scagnostics. <!-- dimensions) where a more complex search is necessary. However, in lower-dimensional problems, the benefit may be less pronounced, as the space is already relatively easy to navigate. While increasing both parameters enhances the performance of JSO, it also extends the computational time required for the optimisation, which can be
computationally intensive when evaluating the index function (such as
scagnostic indexes) multiple times across numerous iterations.-->

@@

@fig-idx-proj-dist shows a sample of traces from the indexes in @tbl-smoothness-squintability for the 6D sine wave data. The traces the PP index values computed for projections along an interpolated path between a noise projection and the optimal projection (sine wave). (Note index values were self-standardized to range between 0 and 1, for comparison.) The `splines` index is very smooth with a gradual increase in value between the noise projection and the optimal projection. The `TIC` index is slightly less smooth and slightly less gradual change in value. Least desirable is the `skinny` index which is noisier and the optimisation needs to be much closer to the optimal angle to have a higher value. This illustrates how the new measures for squintability and smoothness work to describe the PP indexes.

@@

The simulation studies have compared the JSO with CRS and shown how
smoothness and squintability affect the JSO success rate. Here we
explain how they can be used for new index and optimiser development.
Using the JSO optimiser for PPGT requires a change in user behavior.
For all the existing optimisers a single optimisation path is followed.
The JSO has multiple optimisation paths, and needs additional tools to
incorporate into the PPGT, as explained here.

@@

To use JSO for PPGT in the `tourr` package, specify
`search_f = search_jellyfish` in the guided tour. Unlike existing
optimisers, the animation function `animate_*()` won't directly render
the tour path for JSO due to the generation of multiple tour paths. To
visualize the path visited by each individual jellyfish, assign the
animation to an object (`res <- animate_xy(...)`). This will save the
bases visited by JSO, along with relevant metadata, as a tibble data
object [see @RJ-2021-105 for more details on the data object]. The bases
visited for individual jellyfish can then be extracted and viewed using
the `planned_tour()` function.

@@

To sample bases for calculating smoothness, the following arguments are
required: the index function, the dataset, and the number of random
bases to sample. The output of `sample_bases()` is a data frame with a
list-column of sampled basis matrix and index value. Particularization is
available to speed up the index value computation through the
`parallel = TRUE` argument.

@@

This paper has presented new metrics to mathematically define desirable
features of PP indexes, squintability and smoothness, and used these to
assess the performance of the new jellyfish search optimiser. The
metrics will be generally useful for characterizing PP indexes, and help
with developing new indexes.
